---
title: "NGS Sequencing Batch Summary: `r {params$batch_id}`"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    self_contained: true
  tables: yes
params:
  batch_id: NULL
  in_dir: NULL
  in_key: NULL
  in_method: NULL
  in_batch_meta: NULL
  in_config: NULL 
  out_dir: NULL
  ngs_report_version: 1.0.0
---  

# {.tabset}  

## Batch Summary {.active} 
<!-- create details section to collapse multiple no-output code blocks to save space -->
<details style="color: lightgray;"> 
  <summary>Expand Code</summary> 
  
```{r setup, include = TRUE}
# Load Libraries
start_time_all <- Sys.time()

options(warn = 1)  # output warnings as they appear for traceability in stdout/stderr record

knitr::opts_chunk$set(echo = TRUE, warning = FALSE) # warnings will go to console

quiet_library <- function(...) {
  suppressPackageStartupMessages(library(...))
}

quiet_library(batchreporter)
quiet_library(Matrix)        # dependency of H5weaver
quiet_library(rhdf5)         # dependency of H5weaver
quiet_library(H5weaver)      # aifi package
quiet_library(HTOparser)     # aifi package
quiet_library(ggplot2)       
quiet_library(dplyr)         # data wrangling
quiet_library(cowplot)       # arranging multiple plots
quiet_library(gt)            # formatted table output
quiet_library(plotly)        # interactive plots
quiet_library(tidyr)         # data wrangling  
quiet_library(jsonlite)      # reading and writing json metadata files
quiet_library(purrr)         
quiet_library(future)        # multi-threading for batch umap creation
quiet_library(future.apply)  # multi-threading for batch umap creation
quiet_library(Seurat)
library(readr)

stm("Starting NGS Batch Report")

stm(paste(c("\t",paste(names(Sys.info()),Sys.info(),sep = ": ")), collapse = "\n\t"))  
```  

Argument Parsing  
```{r parse_arguments}
# give input directory rna-specific name 
if(is.null(params$in_dir)) {
  batch_id <- "X070"
  in_dir <- system.file("extdata/X070", package = "batchreporter")
  in_key <- system.file("extdata/example_sample_key_X070.csv", package = "batchreporter")
  in_config <- system.file("extdata/default_rna_config_v1.csv", package = "batchreporter")
  in_batch_meta <- system.file("extdata/batch-metadata.json", package = "batchreporter")
  in_method_string <- "scrna;scatac;adt;hto"
  out_dir <- tempdir()
} else {
  batch_id <-  params$batch_id  
  in_method_string <- params$in_method
  in_dir <- params$in_dir  
  in_key <- params$in_key  
  in_config <- params$in_config  
  if(is.null(in_config)){
      in_config <- system.file("extdata/default_rna_config_v1.csv", package = "batchreporter")
      stm("No config file provided. Using default_rna_config_v1.csv for pbmc data")
  }
  out_dir <- params$out_dir
  in_batch_meta <- params$in_batch_meta
  if(is.null(in_batch_meta)){in_batch_meta <- NA}
}

stm(paste0("IN Batch        : ", batch_id))
stm(paste0("IN Method       : ", in_method_string))
stm(paste0("IN Directory    : ", in_dir))
stm(paste0("IN Sample Key   : ", in_key))
stm(paste0("IN Config       : ", in_config))
stm(paste0("IN Batch Processing Info : ", in_batch_meta))
stm(paste0("OUT Dir         : ", out_dir))

print(paste0("IN Batch        : ", batch_id))
print(paste0("IN Method       : ", in_method_string))
print(paste0("IN Directory    : ", in_dir))
print(paste0("IN Sample Key   : ", in_key))
print(paste0("IN Config       : ", in_config))
print(paste0("IN Batch Processing Info : ", in_batch_meta))
print(paste0("OUT Dir         : ", out_dir))
```  

Check input files  
```{r check_input_main}
if(!dir.exists(in_dir)) {
  stm(paste("ERROR: Cannot find IN results dir:", in_dir))
  stop()
}
if(!file.exists(in_key)) {
  stm(paste("ERROR: Cannot find IN sample key:", in_key))
  stop()
}
if(!file.exists(in_config)) {
  stm(paste("ERROR: Cannot find IN config file:", in_config))
  stop()
}
if(!is.na(in_batch_meta) && !file.exists(in_batch_meta)) {
  stm(paste("ERROR: Cannot find IN batch meta:", in_batch_meta))
  stop()
}
if(!dir.exists(out_dir)) {
  stm(paste("Creating output directory:", out_dir))
  dir.create(out_dir)
}

out_prefix <- file.path(out_dir, paste0(batch_id, "_"))

```

Read in the sample key  
```{r read_key}
stm("Reading in sample key")
df_key <- data.table::fread(in_key)
has_controls <- any(grepl("Control", df_key$Type))  # used to control evaluation of batch control-related code chunks

assertthat::assert_that(length(unique(df_key$BatchID)) == 1, msg = "More than 1 batch in input sample key file")
assertthat::assert_that(batch_id == unique(df_key$BatchID), msg = "Batch in input key file does not match input batch value")

```

Determine which modalities streams were run  
```{r determine_modalities}
defined_modalities <- c("scrna", "scatac", "adt", "hto","tcr","bcr")

# convert method string to vector
in_method <- strsplit(in_method_string, split = ";")[[1]]
in_method <- tolower(in_method)

# Logic check input methods
if(!all(in_method %in% defined_modalities)){
    unknowns <- setdiff(in_method, defined_modalities)
    stop(sprintf("One or more input methods are not in defined modalities: '%s'. Defined modalities are: [%s]. Input methods should be passed as a ';'-delimited string, ie 'scrna;scatac;hto'.",
                    paste(unknowns, collapse = "', '"),
                    paste(defined_modalities, collapse = ', ')))
} 

has_rna <- "scrna" %in% in_method
has_atac <- "scatac" %in% in_method
has_adt <- "adt" %in% in_method
has_hto <- "hto" %in% in_method
has_tcr <- "tcr" %in% in_method
has_bcr <- "bcr" %in% in_method

```  

Define and check input folder expectations  
```{r define_input_folders}
if(has_rna){
  in_rna <- file.path(in_dir, "scrna")
  if(!dir.exists(in_rna)){
    stop(sprintf("Expected RNA input directory [%s] does not exist.", in_rna))
  }
}

if(has_atac){
  in_atac <- file.path(in_dir, "scatac")
  if(!dir.exists(in_atac)){
    stop(sprintf("Expected ATAC input directory [%s] does not exist.", in_atac))
  }
}

if(has_adt){
  in_adt <- file.path(in_dir, "adt")
  if(!dir.exists(in_adt)){
    stop(sprintf("Expected ADT input directory [%s] does not exist.", in_adt))
  }
}

if(has_hto){
  in_hto <- file.path(in_dir, "hto")
  if(!dir.exists(in_hto)){
    stop(sprintf("Expected HTO input directory [%s] does not exist.", in_hto))
  }
}

if(has_tcr){
  in_tcr <- file.path(in_dir, "tcr")
  if(!dir.exists(in_tcr)){
    stop(sprintf("Expected tcr input directory [%s] does not exist.", in_tcr))
  }
}


if(has_bcr){
  in_bcr <- file.path(in_dir, "bcr")
  if(!dir.exists(in_bcr)){
    stop(sprintf("Expected bcr input directory [%s] does not exist.", in_bcr))
  }
}

```  

</details> 

### Batch Information    

```{r batchinfo, include = TRUE}
stm("Constructing Batch Information table")

# Summarize batch information, also declare some global batch variables that are used throughout the report
pools <- unique(df_key$PoolID)
n_pools <- length(pools)
if(n_pools==1 & grepl("P0$", pools)){
  n_pools_label <- 0
} else {
  n_pools_label <- n_pools
}

wells <- unlist(strsplit(as.character(unique(df_key$WellID)), split = ";"))
n_wells <- length(wells)

samples <- unique(df_key$SampleID)
n_samples <- length(samples)

controls <- unique(df_key$SampleID[df_key$Type == "Control"])
control_string <- ifelse(has_controls, paste(controls, collapse = ", "), "None")

if (has_controls){
  n_study_samples <- setdiff(samples, controls)
} else {
  n_study_samples <- samples
}

n_study_samples <- length(n_study_samples)

samples_pool <- tapply(df_key$SampleID, df_key$PoolID, unique)
samples_pool_string <- sapply(samples_pool, function(x){paste(x, collapse = ", ")})

labels <- c("Batch","Libraries","N Samples", "N Pools","N Wells","Batch Control", paste0(pools, " Samples"))
values <-  c(batch_id, in_method_string, n_study_samples, n_pools_label, n_wells, control_string, samples_pool_string)

simple_html_table(labels, values, fontsize = 3, col_widths_px = c(175, 850))

```  

### Batch Processing  

```{r display_processing_details, results = "asis"}
if(!is.na(in_batch_meta)){
  if(!file.exists(in_batch_meta)) {stop(sprintf("Supplied in_batch_meta file '%s' does not exist. ",in_batch_meta))}
  meta_table <- jsonlite::read_json(in_batch_meta)
  meta_table <- as.data.frame(meta_table)  # collapses any nested names
  
  if(nrow(meta_table) == 1){
    simple_html_table(colnames(meta_table), unlist(meta_table[1,]), fontsize = 3, col_widths_px = c(200,250))
  } else{
    stop(sprintf("in_batch_meta file should have 1:1 key value pairs. Check supplied file '%s' for proper formatting. ",in_batch_meta))
  }
} else {
  cat("No processing details supplied")
}  
```   



```{r hashing_tab, results = "asis", eval = has_hto, echo = FALSE}
cat("## Cell Hashing", sep = "\n")

```  

```{r hashing_child, child = system.file("rmarkdown/hto_module.Rmd", package = "batchreporter"), eval = has_hto}

```  




```{r scrna_tab, results = "asis", eval = has_rna, echo = FALSE}
cat("## scRNA Well", sep = "\n")

```  

```{r scrna_seq_child, child =  system.file("rmarkdown/scrna_well_module.Rmd", package = "batchreporter"), eval = has_rna}

```   

```{r scrna_hto_tab, results = "asis", eval = has_rna & has_hto, echo = FALSE}
cat("## scRNA Sample", sep = "\n")

```   

```{r hashed_rna_child, child = system.file("rmarkdown/scrna_hashed_sample_module.Rmd", package = "batchreporter"),eval = has_rna & has_hto}

``` 

```{r adt_tab, results = "asis", eval = has_adt, echo = FALSE}
cat("## ADT Well", sep = "\n")

```  

```{r adt_child, child = system.file("rmarkdown/adt_well_module.Rmd", package = "batchreporter"), eval = has_adt}

```   

```{r hashed_adt_tab, results = "asis", eval = has_adt & has_hto, echo = FALSE}
cat("## ADT Sample", sep = "\n")

```

```{r hashed_adt_child, child = system.file("rmarkdown/adt_hashed_sample_module.Rmd", package = "batchreporter"), eval = has_adt & has_hto}

```



```{r atac_tab, results = "asis", eval = has_atac, echo = FALSE}
cat("## scATAC Well", sep = "\n")

```  

```{r atac_child, child =  system.file("rmarkdown/scatac_seq_module.Rmd", package = "batchreporter"), eval =  has_atac}

```   

```{r hashed_atac_tab, results = "asis", eval = has_atac & has_hto, echo = FALSE}
cat("## scATAC Sample", sep = "\n")

```  
  
```{r hashed_atac_child, child = system.file("rmarkdown/scatac_hashing_module.Rmd", package = "batchreporter"), eval = has_atac & has_hto}

```   

 
```{r tcr_tab, results = "asis", eval = has_tcr, echo = FALSE}
cat("## scTCR Well", sep = "\n")

```  

```{r tcr_child, child =  system.file("rmarkdown/tcr_well_module.Rmd", package = "batchreporter"), eval = has_tcr}

```  


```{r hashed_tcr_tab, results = "asis", eval = has_tcr & has_hto, echo = FALSE}
cat("## scTCR Sample", sep = "\n")

```  
  
```{r hashed_tcr_child, child = system.file("rmarkdown/tcr_hashed_sample_module.Rmd", package = "batchreporter"), eval = has_tcr & has_hto}


```{r bcr_tab, results = "asis", eval = has_bcr, echo = FALSE}
cat("## scBCR Well", sep = "\n")

```  

```{r bcr_child, child =  system.file("rmarkdown/bcr_well_module.Rmd", package = "batchreporter"), eval = has_bcr}

```  

```{r hashed_bcr_tab, results = "asis", eval = has_bcr & has_hto, echo = FALSE}
cat("## scBCR Sample", sep = "\n")

```  

```{r hashed_bcr_child, child =  system.file("rmarkdown/bcr_hashed_sample_module.Rmd", package = "batchreporter"), eval = has_bcr & has_hto}

```  
  
  
## Session Information   


<a id="top_session_info"></a>  

Input Directory:  

```{r echo_in_dir}
in_dir 

```  
Input Directory Contents:  

```{r echo_in_files}
folders <- list.dirs(in_dir, recursive = FALSE)

file_list <- lapply(folders, function(x){
  dir(x, recursive = TRUE)
})
names(file_list) <- basename(folders)
file_list

```  

Batch processing metadata:  

```{r echo_processing_method}
in_batch_meta 

```  

Config File:  
```{r echo_config_files}
in_config

```  

Key File:  
```{r echo_key_files}
in_key

```  

Output Directory:  
```{r echo_out_dir}
out_dir

``` 

Session Info:  
```{r session_info_main}
sessionInfo()
```

Total time elapsed
```{r show_time_main}
end_time_all <- Sys.time()
diff_time_all <- end_time_all - start_time_all
time_message_all <- paste0("Elapsed Time: ", 
                       round(diff_time_all, 3),
                       " ", units(diff_time_all))
print(time_message_all)
stm(time_message_all)
stm("Batch report process complete.")
```  

[Return to Top](#top_session_info)

# {-}

---  

NGS report v.`r {params$ngs_report_version}`, Lauren Okada
